Andres Karjus

### 2020-02-13 index.Rmd
news

<div style="text-align: justify; padding-left:1px"> 
I am a PhD student in linguistics at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/), working on lexical dynamics and evolutionary variant selection in language with [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/), utilizing massive corpora and natural language processing to hunt traces of selection and competition in language. This research is supported by a Kristjan Jaak scholarship from the Archimedes Foundation of Estonia. I am also affiliated as a (part-time) junior researcher with the University of Tartu [EKKAM](http://ekkam.ut.ee/en/) sociolinguistics group, doing data analysis and agent-based models. I am also engaged in teaching and consulting (see the tab above) stats and data visualization for the humanities and social sciences. On the right is the picture obligatory for all academic home pages to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2018. Challenges in detecting evolutionary forces in language change using diachronic corpora | [arXiv preprint](http://arxiv.org/abs/1811.01275) | [code](https://github.com/andreskarjus/wfsim_fit) | in peer review
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Quantifying the dynamics of topical fluctuations in language. | [arXiv preprint](https://arxiv.org/abs/1806.00699) | <a href="karjus_quantifying_2018.bib" target="_blank">bib</a> | [code](https://github.com/andreskarjus/topical_cultural_advection_model) | in press at Language Dynamics and Change
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [Online at Brill](https://doi.org/10.1163/22105832-00802004) | [postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [online version](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>

@@

Before coming to Edinburgh, I worked as a teaching assistant in informatics at the University of Tartu, before that studied artificial intelligence and natural language processing at KU Leuven (MSc) and linguistics at the University of Tartu (BA, MA). Somewhere in between I also did exchanges to the University of Iceland and the University Vienna, attended a dozen-odd academic summer schools, taught Icelandic to art students and Estonian to Norwegian teachers, worked as an assistant at the Estonian Wordnet project, and did internships at CrossLang NV in Belgium and at the (then still existing linguistics department of the) Max Planck Institute for Evolutionary Anthropology in Leipzig. In earlier years, I worked various studenty sort of jobs to support my studies (not to list them all, but one of them involved swords and another one regularly mowing grass off the roof of a small yellow building).
</div>
<br>

>>>>

<div style="text-align: justify; padding-left:1px"> 
I am a PhD student in linguistics at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/), working on lexical dynamics and evolutionary variant selection in language with [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/), utilizing massive corpora and natural language processing to hunt traces of selection and competition in language; and experiments to see how individuals perform given limited linguistic resources. This research is supported by a Kristjan Jaak scholarship from the Archimedes Foundation of Estonia. I am also engaged in teaching and consulting (see the tab above) stats and data visualization for the humanities and social sciences. On the right is the picture obligatory for all academic home pages to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2018. Challenges in detecting evolutionary forces in language change using diachronic corpora | [arXiv preprint](http://arxiv.org/abs/1811.01275) | [code](https://github.com/andreskarjus/wfsim_fit) | in press at Glossa
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. [Language Dynamics and Change](https://doi.org/10.1163/22105832-01001200) | [arXiv preprint](https://arxiv.org/abs/1806.00699) | <a href="" target="_blank">bib</a> | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [Online at Brill](https://doi.org/10.1163/22105832-00802004) | [postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [online version](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>

@@

Before coming to Edinburgh, I worked as a teaching assistant in informatics at the University of Tartu (2015-2016), before that studied artificial intelligence and natural language processing at KU Leuven (MSc) and linguistics at the University of Tartu (BA, MA). I was also affiliated 2016-2019 as a (part-time) junior researcher with the University of Tartu [EKKAM](http://ekkam.ut.ee/en/) sociolinguistics group, doing data analysis and agent-based models. 
During my pre-PhD studies I also went on exchanges to the University of Iceland and the University Vienna, attended a dozen-odd academic summer schools, taught Icelandic to art students and Estonian to Norwegian teachers, worked as an assistant at the Estonian Wordnet project, and did internships at CrossLang NV in Belgium and at the (then still existing linguistics department of the) Max Planck Institute for Evolutionary Anthropology in Leipzig. In earlier years, I worked various studenty sort of jobs to support my studies (not to list them all, but one of them involved swords and another one regularly mowing grass off the roof of a small yellow building).
</div>
<br>

### 2020-02-13 news.Rmd
news

>>>>

- My first PhD paper (that's been out as a preprint since summer 2018) is now published in Language Dynamics and Change: Karjus, A., Blythe, R. A., Kirby, S., & Smith, K. (2020). Quantifying the dynamics of topical fluctuations in language. Language Dynamics and Change. https://doi.org/10.1163/22105832-01001200

- Delivered another dataviz workshop, this time for the University of Edinburgh Centre for Data, Culture and Society (01.11.2019)
